### 2018-01-04 UOW_thesis_template_v2-0/thesis_template/chapters/Intro.tex
:wrench: Removing the blue text after Marianito's approval

	A well-known result for determining the option value is known as the Black-Scholes equation \cite{Black1973}. The Black-Scholes formula is in fact a partial differential equation (PDE) that is also paired with a terminal condition governed by a payoff function. The payoff function merely states what the option's value will be at the expiry date. Typically, this is normally a piecewise linear function of the strike price and the underlying asset (at least in terms of a call and put). One of the most classical methods to solve the Black-Scholes PDE system is to invoke a transformation of variables (further details of this in addition to~\cite{Black1973} can be found in~\cite{Wilmott1995}) that effectively eliminates a few terms and reduces the PDE system to the archetypal heat equation commonly seen in thermodynamics~\cite{cannon1984one}. The motivation for this step is because the heat equation has been studied endlessly in literature and the solution to the PDE system is ascertainable~\cite{Churchill1938}. The original details of~\cite{Black1973} can also be complemented with~\cite{Wilmott1995} for further details and discussion about this derivation. The crux of option valuation is to determine a fair price for the holder to pay the writer to enter into their specified contract. This price is commonly known as the \emph{option premium} (or simply, the premium) that actually signifies the value of the option at time-zero (i.e., starting today).

@@

Financially, American option contracts are more ideal to trade because one can exercise the option at any time up until and including the expiry date. Perhaps the earliest mathematical formulation for the American option pricing problem was proposed by McKean~\cite{McKean1965} where it was proved that there is an equivalence between the optimal stopping problem describing the framework of the American option valuation problem and a free boundary problem. McKean was then able to derive a homogeneous PDE on a restricted domain, but this was solved using an incomplete Fourier transform and an analytical result was obtained. Geske~\cite{Geske1984} then employed a discrete time approach using a financial derivative called compound options (this will be discussed later in subsection 1.2). The basic idea Geske employed was to first discretize the time domain then price an American put option as the discounted expected value of all future cash flows. The justification for these cash flows is because the put option can be exercised at any discrete interval in time for the duration of the option's lifespan. Another perspective to this is decomposing the American option into a finite quantity of European style options. It is then illustrated that this construction is valid due to it being a solution of the Black-Scholes PDE subject to the free boundary condition imposed by the presence of the optimal exercise boundary.

@@

Geske~\cite{Geske1979} developed the seminal theory for compound option models. The motivation was to be able to accurately price a firm's common stock (also known as ordinary share) -- a security that grants the holder corporate equity ownership. Black and Scholes~\cite{Black1973} argued that common stock can be interpreted as an option on the firm. This is because when a firm defaults, the holders of the common stock maintain the right but not necessarily the obligation to sell the entire firm to the bondholders (who possess claims to the firm's future cash flows as a result of its financial liabilities) for a strike equal to the face value of the bond \cite{Kwok2008}. Thus an option on a portion of the common stock can be viewed as a compound option since the underlying received upon exercising the option is technically another option on the value of the firm. Consequently, Geske~\cite{Geske1979} derived analytical formulas for compound options when the firm value follows a geometric Brownian motion. This work was then extended to price compound options on various types of bonds (e.g., risky coupon bonds~\cite{Geske1977}, retractable and extendible bonds~\cite{Brennan1977, Ananthanarayanan1980, Longstaff1990}).

@@

\section{Jump-diffusion models}
	It was verified by Merton \cite{Merton1973b} that one of the fundamental assumptions of the Black-Scholes model is that the asset price follows a continuous-time, diffusion process with a continuous sample path. This prompted Merton in \cite{Merton1976} to consider a ``jump'' stochastic process for the asset price that allows for the probability for it to change at large magnitudes irrespective of the time interval between successive observations. The jumps in the asset price can be accommodated by appending an additional source of uncertainty into the asset price dynamics that models the discontinuity. Moreover, subsequent empirical studies (e.g., Rosenfeld \cite{Rosenfeld1980}, Jarrow \& Rosenfeld~\cite{Jarrow1984}, Ball \& Torous \cite{Ball1985}, and Brown \& Dybvig \cite{Brown1986}) asserted that the asset price process is best modelled by a stochastic process with a discontinuous sample path. This phenomenon suggests that the asset price dynamics follow a  \emph{jump-diffusion model}.

@@

The importance of developing such a system extends beyond attempting to capture the options market's behaviour at any given point. The need lies within being able to deliver fundamental explanations to why certain phenomena occur. For example, when one wishes to estimate the implied volatility surfaces to calibrate the standard Black-Scholes option values to actual market quotes, the Black-Scholes model where the underlying asset follows a standard diffusion process assumes the implied volatility surface to be flat. That is, a constant value during the option's lifetime and for varying values of the strike price (options are commonly listed as a function of their strike price). But empirical observations have shown that these implied volatility surfaces are heavily dependent on both the strike price and the expiry date (in particular, refer to Heynen~\cite{Heynen1994}, Dumas \emph{et al.}~\cite{Dumas1998}, Rebonato~\cite{Rebonato1999}, and Cont \& Fonseca~\cite{Cont2001, Cont2002}). As a result, these surfaces actually form either a ``smile'' or ``skew'' depending on the values of the strike and time to expiry. Dupire~\cite{Dupire1994} developed a technique for computing the local implied volatility surfaces and he showed that the standard Black-Scholes model with an asset under diffusion dynamics can embody all the distinguishing features of this ``smile problem''. However, it only gives us a tool needed to ensure we recover the required option values. It does not explain why these smiles and skews occur. A jump-diffusion model, however, is able to encapsulate both a justification for these smiles and skews, their increased occurrences after the 1987 crash (see Andersen and Andreasen~\cite{Andersen2000}), and how the jumps in the asset price reflect the "jump fear" in market participants~\cite{Cont2004}.

@@

Frontczak \cite{Frontczak2013} adopted a method of solving the PIDE seen in \cite{Merton1976} using Mellin transforms. He proceeded to re-derive Merton's solution for a European put option via direct inversion. Frontczak's approach of directly evaluating the inverse Mellin integral (i.e., a complex integral) \textcolor{black}{is where the approach could be improved}. Moreover, this process needs to be repeated for different payoffs, making this procedure computationally expensive and tedious.

\section{Implied volatility}
		In the Black-Scholes option pricing model, most of the associated parameters (e.g., the option price, interest rate) are observable. The only quantity that cannot be observed is the variable $\sigma$: the volatility. Mathematically, it is significant as its role is to emulate a level of uncertainty in the underlying asset. In practice, its importance is further signified because prior knowledge of $\sigma$ would enable a financial practitioner to accurately price other derivatives that also incorporate the same underlying asset they are interested in. 
		For estimating the volatility~$\sigma$ in the standard diffusion model~\eqref{eqn:stock}, there exist two primary methods. The first scheme involves estimating $\sigma$ from previous asset price movements. That is, suppose a model for the behaviour of the asset involving $\sigma$ is known and the asset prices for all times up until the present are accessible. Then $\sigma$ can be fitted to this observed data. This method is dubbed \emph{historical volatility} as $\sigma$ is approximated using data of previous asset prices. The second is to calibrate all the known parameters, then treat the option value as a function of $\sigma$ (the option price is one of the known parameter values) and solve for $\sigma$. This approach determines $\sigma$ implicitly from the Black-Scholes formula using the option price and the observed parameters, and is referred to as \emph{implied volatility}.  Aside from option pricing in a jump-diffusion framework, another aim of this article is to present a novel implied volatility scheme using Mellin transforms. For the scope of this thesis, we will be limiting the discussions of implied volatility to options where the \emph{vega} (i.e., a measure of sensitivity of options prices to changes in volatility~\cite{Vine2011}) has only one sign. This assumption of a single-signed vega includes European calls and puts.

@@

\section{Outline of the thesis}
		The thesis will be structured as follows. All the necessary preliminary knowledge will be contained in chapter 2. This will cover the introductory content for the standard European and American options as well as options on underlying assets that are subjected to jump-diffusion dynamics, and the Mellin transform. These will form the basis for all the work to come.

@@

		In chapter 6, we propose a method for computing European compound options with general payoffs assuming the underlying asset follows a geometric Brownian motion. We also show that for standard vanilla European options, we only need to generate two cases; the remaining two can be ascertained via a \emph{pseudo-put-call parity} technique. We can also obtain a combined result that resembles the standard put-call parity identity for compound options, and this stems from using the pseudo-put-call parity approach to derive the four standard vanilla compound option cases. We then extend to pricing European compound options when the underlying pays one discrete dividend. Although analytical formulas for European compound options with continuous dividend yields for the underlying have been derived before in the literature, to our knowledge this has not been done for discrete paying dividends either as a cash payment or proportional yield of the underlying asset. The approach we adopt in this work incorporates the Black-Scholes kernel and the integral identities associated with it to give an exact pricing formula for compound options on a underlying that has a continuous dividend yield. With a slight modification, we also show that this approach is valid when the underlying pays one discrete dividend. We will model the dynamics of the discrete dividend payment as a yield of the underlying asset.

>>>>

	A well-known result for determining the option value is known as the \textcolor{black}{Black-Scholes equation} \cite{Black1973}. The Black-Scholes formula is in fact a partial differential equation (PDE) that is also paired with a terminal condition governed by a payoff function. The payoff function merely states what the option's value will be at the expiry date. Typically, this is normally a piecewise linear function of the strike price and the underlying asset (at least in terms of a call and put). One of the most classical methods to solve the Black-Scholes PDE system is to invoke a transformation of variables (further details of this in addition to~\cite{Black1973} can be found in~\cite{Wilmott1995}) that effectively eliminates a few terms and reduces the PDE system to the archetypal heat equation commonly seen in thermodynamics~\cite{cannon1984one}. The motivation for this step is because the heat equation has been studied endlessly in literature and the solution to the PDE system is ascertainable~\cite{Churchill1938}. The original details of~\cite{Black1973} can also be complemented with~\cite{Wilmott1995} for further details and discussion about this derivation. The crux of option valuation is to determine a fair price for the holder to pay the writer to enter into their specified contract. This price is commonly known as the \emph{option premium} (or simply, the premium) that actually signifies the value of the option at time-zero (i.e., starting today).

@@

\textcolor{black}{Financially, American option contracts are more ideal to trade because one can exercise the option at any time up until and including the expiry date}. Perhaps the earliest mathematical formulation for the American option pricing problem was proposed by McKean~\cite{McKean1965} where it was proved that there is an equivalence between the optimal stopping problem describing the framework of the American option valuation problem and a free boundary problem. McKean was then able to derive a homogeneous PDE on a restricted domain, but this was solved using an incomplete Fourier transform and an analytical result was obtained. Geske~\cite{Geske1984} then employed a discrete time approach using a financial derivative called compound options (this will be discussed later in subsection 1.2). The basic idea Geske employed was to first discretize the time domain then price an American put option as the discounted expected value of all future cash flows. The justification for these cash flows is because the put option can be exercised at any discrete interval in time for the duration of the option's lifespan. Another perspective to this is decomposing the American option into a finite quantity of European style options. It is then illustrated that this construction is valid due to it being a solution of the Black-Scholes PDE subject to the free boundary condition imposed by the presence of the optimal exercise boundary.

@@

Geske~\cite{Geske1979} developed the seminal theory for compound option models. The motivation was to be able to accurately price a firm's common stock (also known as ordinary share) -- a security that grants the holder corporate equity ownership. Black and Scholes~\cite{Black1973} argued that common stock can be interpreted as an option on the firm. \textcolor{black}{This is because when a firm defaults}, the holders of the common stock maintain the right but not necessarily the obligation to sell the entire firm to the bondholders (who possess claims to the firm's future cash flows as a result of its financial liabilities) for a strike equal to the face value of the bond \cite{Kwok2008}. Thus an option on a portion of the common stock can be viewed as a compound option since the underlying received upon exercising the option is technically another option on the value of the firm. Consequently, Geske~\cite{Geske1979} derived analytical formulas for compound options when the firm value follows a geometric Brownian motion. This work was then extended to price compound options on various types of bonds (e.g., risky coupon bonds~\cite{Geske1977}, retractable and extendible bonds~\cite{Brennan1977, Ananthanarayanan1980, Longstaff1990}).

@@

\section{Jump-diffusion models}
	It was verified by Merton \cite{Merton1973b} that one of the fundamental assumptions of the Black-Scholes model is that the asset price follows a continuous-time, diffusion process with a continuous sample path. This prompted Merton in \cite{Merton1976} to consider a ``jump'' stochastic process for the asset price that allows for the probability for it to change at large magnitudes irrespective of the time interval between successive observations. The jumps in the asset price can be accommodated by appending an additional source of uncertainty into the asset price dynamics that models the discontinuity. Moreover, subsequent empirical studies (e.g., Rosenfeld \cite{Rosenfeld1980}, Jarrow \& Rosenfeld~\cite{Jarrow1984}, Ball \& Torous \cite{Ball1985}, and Brown \& Dybvig \cite{Brown1986}) \textcolor{black}{asserted that the asset price process is best modelled by a stochastic process with a discontinuous sample path}. This phenomenon suggests that the asset price dynamics follow a  \emph{jump-diffusion model}.

@@

The importance of developing such a system extends beyond attempting to capture the options market's behaviour at any given point. The need lies within being able to deliver fundamental explanations to why certain phenomena occur. For example, when one wishes to estimate the implied volatility surfaces to calibrate the standard Black-Scholes option values to actual market quotes, the Black-Scholes model where the underlying asset follows a standard diffusion process assumes the implied volatility surface to be flat. That is, a constant value during the option's lifetime and for varying values of the strike price (options are commonly listed as a function of their strike price). But empirical observations have shown that these implied volatility surfaces are heavily dependent on both the strike price and the expiry date (in particular, refer to Heynen~\cite{Heynen1994}, Dumas \emph{et al.}~\cite{Dumas1998}, Rebonato~\cite{Rebonato1999}, and Cont \& Fonseca~\cite{Cont2001, Cont2002}). As a result, these surfaces actually form either a ``smile'' or ``skew'' depending on the values of the strike and time to expiry. Dupire~\cite{Dupire1994} developed a technique for computing the local implied volatility surfaces and he showed that the standard Black-Scholes model with an asset under diffusion dynamics can embody all the distinguishing features of this ``smile problem''. However, it only gives us a tool needed to ensure we recover the required option values. It does not explain why these smiles and skews occur. A jump-diffusion model, however, is able to encapsulate both a justification for these smiles and skews, their increased occurrences after the 1987 crash (see Andersen and Andreasen~\cite{Andersen2000}), \textcolor{black}{and how the jumps in the asset price reflect the "jump fear" in market participants~\cite{Cont2004}}.

@@

Frontczak \cite{Frontczak2013} adopted a method of solving the PIDE seen in \cite{Merton1976} using Mellin transforms. He proceeded to re-derive Merton's solution for a European put option via direct inversion. Frontczak's approach of \textcolor{black}{directly evaluating the inverse Mellin integral (i.e., a complex integral)} \textcolor{black}{is where the approach could be improved}. Moreover, this process needs to be repeated for different payoffs, making this procedure computationally expensive and tedious.

\section{Implied volatility}
		In the Black-Scholes option pricing model, \textcolor{black}{most of the associated parameters (e.g., the option price, interest rate) are observable}. The only quantity that cannot be observed is the variable $\sigma$: the volatility. Mathematically, it is significant as its role is to emulate a level of uncertainty in the underlying asset. In practice, its importance is further signified because prior knowledge of $\sigma$ would enable a financial practitioner to accurately price other derivatives that also incorporate the same underlying asset they are interested in. 
		For estimating the volatility~$\sigma$ in the standard diffusion model~\eqref{eqn:stock}, there exist two primary methods. The first scheme involves estimating $\sigma$ from previous asset price movements. That is, suppose a model for the behaviour of the asset involving $\sigma$ is known and the asset prices for all times up until the present are accessible. Then $\sigma$ can be fitted to this observed data. This method is dubbed \emph{historical volatility} as $\sigma$ is approximated using data of previous asset prices. The second is to calibrate all the known parameters, then treat the option value as a function of $\sigma$ (the option price is one of the known parameter values) and solve for $\sigma$. This approach determines $\sigma$ implicitly from the Black-Scholes formula using the option price and the observed parameters, and is referred to as \emph{implied volatility}.  Aside from option pricing in a jump-diffusion framework, another aim of this article is to present a novel implied volatility scheme using Mellin transforms. For the scope of this thesis, we will be limiting the discussions of implied volatility to options where the \emph{vega} (i.e., a measure of sensitivity of options prices to changes in volatility~\cite{Vine2011}) has only one sign. This assumption of a single-signed vega includes European calls and puts.

@@

\section{Outline of the thesis}
		The thesis will be structured as follows. All the necessary preliminary knowledge will be contained in chapter 2. This will cover the introductory content for the standard European and American options as well as options on underlying assets that are subjected to jump-diffusion dynamics, and the Mellin transform. These will form the \textcolor{black}{basis} for all the work to come.

@@

		In chapter 6, we propose a method for computing European compound options with general payoffs assuming the underlying asset follows a geometric Brownian motion. We also show that for standard vanilla European options, we only need to generate two cases; the remaining two can be ascertained via a \emph{pseudo-put-call parity} technique. We can also obtain a combined result that resembles the standard put-call parity identity for compound options, and this stems from using the pseudo-put-call parity approach to derive the four standard vanilla compound option cases. We then extend to pricing European compound options when the underlying pays one discrete dividend. \textcolor{black}{Although} analytical formulas for European compound options with continuous dividend yields for the underlying have been derived before in the literature, to our knowledge this has not been done for discrete paying dividends either as a cash payment or proportional yield of the underlying asset. The approach we adopt in this work incorporates the Black-Scholes kernel and the integral identities associated with it to give an exact pricing formula for compound options on a underlying that has a continuous dividend yield. With a slight modification, we also show that this approach is valid when the underlying pays one discrete dividend. We will model the dynamics of the discrete dividend payment as a yield of the underlying asset.
